# Before we begin

Make sure you have the tidyverse instaled:

```{r, eval = FALSE}
install.packages("tidyverse")
install.packages("gapminder")
usethis::use_course("http://bit.ly/30lAoSe")
```

# Welcome

My name is Hadley. My job today is to help boost your tidyverse productivity:

* Thinking about types
* Using functional programming
* Everything you really need to know about tidy eval

And I'm going to be teaching you all this with live coding, which (full disclosure), I've never done before! 

Your job:

* Use post-its to give me feedback throughout the workshop (green = good; pink = slow down/problem)
* Please ask questions! (Either to whole class or to me individually during breaks).
  In particularly, if I do something that seems particularly magical 
  (i.e. I don't type enough for what you see on the screen), tell me!

At the end of the day I'll post all the code I type at <https://github.com/hadley/tidyverse-booster>. Avoid looking at it before then.

## Warm ups

Introduce yourself to your neighbours. What is your favourite dplyr function and why?

A few challenges to get your brain warmed up:

* What tidyverse packages are design to help with strings, date-times, factors?
  What does the glue package do?

* What does this code do?

```{r}
library(tidyverse)

x <- 1:40
case_when(
  x %% 15 == 0 ~ "fizz buzz",
  x %%  5 == 0 ~ "buzz",
  x %%  3 == 0 ~ "fizz",
  TRUE         ~ as.character(x)
)
```

```{r}
x <- c(1, NA, 3, NA, NA)
y <- c(3, 2,  9, 4, NA)
z <- c(5, 4, NA, 10, 5)

coalesce(x, y, z)
```

```{r}
# https://apps.garrickadenbuie.com/countdown/
countdown::countdown_app()
```

