# Functional programming

<!-- Create new file -->

Big topic: loosely refers to functions that have a function as input or output.

```{r}
integrate(sin, 0, 2 * pi)

x <- c(rnorm(20, mean = 0, sd = 1), rnorm(30, mean = 0, sd = 2))
f <- ecdf(x)
f(0)
f(3)
plot(f)
```

Today we're going to focus on the `map()` family of functions from purrr, because they have a big pay off, typically for data import and report generation. Specifically focus on making reports.

They're an alternative to a common class of for loop which basically looks like:

```{r}
for (i in seq_along(input)) {
  output[[i]] <- f(input[[i]])
}
```

i.e. the element i of the output only depends on element i of the input

## Why not for loops?

Interlude with slides

## Data import

```{r}
usethis::use_course("http://bit.ly/30lAoSe")
```

* Read `excel/gapminder-country.xls`
* Read `csv/country/*.csv`

Your turn: repeat for year. Repeat for continent.

## Type-stability

`sapply()`

```{r}
dt <- data.frame(
  x = ordered("a"),
  y = Sys.time(),
  z = 1
)
sapply(dt[1:3], class)
sapply(dt[1:2], class)
sapply(dt[3], class)
```

`vapply()` is type-stable, but is a lot of typing

## Parameterised R markdown

Open `report-country.Rmd` and knit it.  Where does the country come from? Can you make it display data for the United Kingdom instead?

Read the help for `rmarkdown::render()`. Can you make it render separate PDFs for New Zealand and the United Kingdom? (Without touching the .Rmd)

Can you write a for-loop that renders each country to a separate report?

`walk2()` is a wrapper for the following for-loop:

```{r}
# walk2(x, y, ~ f(.x, .y))

for (i in 1:length(x)) {
  f(x[[i]], y[[i]])
}
```

Can you rewrite the for loop to use `walk2()` instead?

<!-- COMMIT TO GIT -->
