# tidyeval

<!-- Create new file -->

Tidy evaluation is the system used throughout the tidyverse that allows you to type variable names without typing the data frame name:

```{r}
library(tidyverse)

diamonds %>% 
  ggplot(aes(carat, price)) + 
  geom_hex()

diamonds %>% 
  group_by(color, clarity) %>% 
  summarise(price = mean(price))

carat
price
color
clarity
```

We call this __data masking__. Helps typing, and makes tools like dbplyr and dtplyr possible.

Generally, when you're doing a data analysis, this just works without you having to think about it. But at some point you're going to produce a bunch of duplicated code with copy and paste:

```{r}
diamonds %>% 
  group_by(color) %>% 
  summarise(price = mean(price))
diamonds %>% 
  group_by(clarity) %>% 
  summarise(price = mean(price))
diamonds %>% 
  group_by(cut) %>% 
  summarise(price = mean(price))
```

And you'll want to make a function. If you use your existing function making skills, you'll hit a problem:

```{r}
mean_price <- function(df, by) {
  diamonds %>% 
    group_by(by) %>% 
    summarise(price = mean(price))
}
diamonds %>% mean_price(by = cut)
```

Now rather than referring to the column directly, you want to refer to it indirectly - you want to tell `group_by` to group by the varible recorded in `by`, not literally group by the variable called `by`.

Fixing this problem requires two steps:

1. Identify that it's a tidyeval problem; i.e. you're trying to refer to
   a data frame variable that's stored in a function argument; and you're
   inside the tidyverse

2. Know about embracing `{{ }}`

```{r}
mean_price <- function(df, by) {
  diamonds %>% 
    group_by({{ by }}) %>% 
    summarise(price = mean(price))
}
diamonds %>% mean_price(by = cut)
diamonds %>% mean_price(by = color)
```

Inspired by the very awesome glue package:

```{r}
library(glue)
name <- "Hadley"
glue("Hello {name}!")
```


(Previously you had to know a bunch of theory about how tidyeval works; that's no longer the case!)

Seriously that's now 90% of what you need to know to use tidyeval for simple cases.

---

```{r}
library(tidyverse)

mtcars %>% pull(am)

by_cyl <- mtcars %>%
  group_by(cyl) %>%
  summarise(mean = mean(mpg))

ggplot(by_cyl, aes(cyl, mpg)) + 
  geom_point()
```


## Other places where you don't need tidy eval

### Data

```{r}
diamonds %>% 
  filter(x == 0) %>% 
  summarise(
    n = n(), 
    price = paste0(min(price), "--", max(price)),
    carat = paste0(min(carat), "--", max(carat)),
  )

diamonds %>% 
  filter(y == 0) %>% 
  summarise(
    n = n(), 
    price = paste0(min(price), "--", max(price)),
    carat = paste0(min(carat), "--", max(carat)),
  )

diamonds %>% 
  filter(z == 0) %>% 
  summarise(
    n = n(), 
    price = paste0(min(price), "--", max(price)),
    carat = paste0(min(carat), "--", max(carat)),
  )
```

How can we reduce the duplication here? Your turn: first make a range summary function.

```{r}
range_sum <- function(x) {
  
}

# Bonus points: optionally allow specification of the number decimal places too
```

Next make a diamonds summary function:

```{r}
diamonds_sum <- function(df) {
  df %>% 
    summarise(
      n = n(),
      price = range_sum(price),
      carat = range_sum(carat)
    )
}
```

This is easy because the data frame argument doesn't use tidy evaluation! You should always check that tidy eval is needed before you try something more complicated.

### Dots

Another case that lets you avoid tidy eval is taking advantage of `...` to forward on multiple arguments.

For example, we could rewrite `mean_price()`  using `...`:

```{r}
mean_price <- function(df, ...) {
  diamonds %>% 
    group_by(...) %>% 
    summarise(price = mean(price))
}
diamonds %>% mean_price(cut, color)
```

The advantage is that we can group by any number of variables; the disadvantage is that it becomes harder to pass on other arguments (and `...` silently swallows misnamed args)


## Other challenges

Setting names requires `:=`.

```{r}
count_top <- function(df, var, n = 5) {
  df %>%
    mutate({{ var }} := fct_lump(fct_infreq({{ var }}), n = n)) %>%
    group_by({{ var }}) %>%
    summarise(n = as.integer(sum(weight)))
}
```

<!-- COMMIT TO GIT -->
